---
title: "Binary matrix for confidence scoring"
author: "Cosette Schupfer"
date: "04/08/2020"
output: html_document
---

Creating this matrix allowed calculation of the confidence scores for each of the genes. 
#load data
PAXFOXG1TARGETS is the list of 396 common genes established from overlapping the lists from 5 datasets. ALLDATASETS is a csv file containing all gene lists, separated by columns.
```{r}
PAX6FOXG1Targets <- read.csv("C:/Users/user/OneDrive/Documents/R/Confidence score/Foxg1Pax6Gene_Overlap.csv")
ALLDATASETS <- read.csv("C:/Users/user/OneDrive/Documents/R/Confidence score/CFGENELIST.csv")
```
#Match the genes in PAX6FOXG1Targets to the genes in ALLDATASETS
If the gene is features, the matrix returns 1, if not it returns 0. This facilitates simple summation of the scores to calculate the final confidence score
```{r}
PAX6FOXG1Targets$XPax6_E15.5F = 0
PAX6FOXG1Targets$XPax6_E15.5F[PAX6FOXG1Targets$GENE %in% ALLDATASETS$ï..XPax6_E15.5F] <- 1

PAX6FOXG1Targets$SPax6_E12.5NC = 0
PAX6FOXG1Targets$SPax6_E12.5NC[PAX6FOXG1Targets$GENE %in% ALLDATASETS$SPax6_E12.5NC] <- 1

PAX6FOXG1Targets$QU_Foxg1cKOvsWT = 0
PAX6FOXG1Targets$QU_Foxg1cKOvsWT[PAX6FOXG1Targets$GENE %in% ALLDATASETS$QU_Foxg1cKOvsWT] <- 1

PAX6FOXG1Targets$F_Foxg1HOMvsWT = 0
PAX6FOXG1Targets$F_Foxg1HOMvsWT[PAX6FOXG1Targets$GENE %in% ALLDATASETS$F_Foxg1HOMvsWT] <- 1

PAX6FOXG1Targets$N_Pax6KOvsWT = 0
PAX6FOXG1Targets$N_Pax6KOvsWT[PAX6FOXG1Targets$GENE %in% ALLDATASETS$N_Pax6KOvsWT] <- 1

PAX6FOXG1Targets$G_Pax6HOMvsWT = 0
PAX6FOXG1Targets$G_Pax6HOMvsWT[PAX6FOXG1Targets$GENE %in% ALLDATASETS$G_Pax6HOMvsWT] <- 1

PAX6FOXG1Targets$QU_Pax6cKOvsWT = 0
PAX6FOXG1Targets$QU_Pax6cKOvsWT[PAX6FOXG1Targets$GENE %in% ALLDATASETS$QU_Pax6cKOvsWT] <- 1
```
#Write to file to get a csv file output
```{r}
#write to file
write.csv(PAX6FOXG1Targets, "C:/Users/user/OneDrive/Documents/R/Confidence score/GeneBinaryMatrix.csv")
```

