---
title: "GO analysis"
author: "Cosette Schupfer"
date: "04/08/2020"
output: html_document
---
Gene Ontology analysis. Note that this analysis was based on results from ChIP experiments therefore fold change (denoting direction of differential expression) was replaced with confidence score. FDR was set to 0 as a default. 
#load packages
```{r, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(DOSE)
library(GO.db)
library(org.Mm.eg.db)
library(clusterProfiler)
library(goseq)
library(ggplot2)
library(enrichplot)
library(biomaRt)
library(kableExtra)
library(knitr)


```
#Import data

```{r}
df <- read.csv("C:/Users/user/OneDrive/Documents/R/GO analysis/GENELIST.csv")
colnames(df) <- c("Gene","FDR","FC")
```

#define parameters
```{r}
gene_ID = df$Gene
FDR_cutoff = 0.05
FC_cutoff = 0
organism.GO = "org.Mm.eg.db"
organism.biomart = "mmusculus_gene_ensembl"
ontology = "BP"
```

#convert Gene ID to EntrezID
```{r, warning=FALSE}
entrez_genes = bitr(df$Gene, fromType="SYMBOL", toType="ENTREZID", OrgDb=organism.GO)
df <- merge(df, entrez_genes, by.x="Gene", by.y="SYMBOL")
```

genes_up actually denotes all of the genes as a confidence score cut-off of 0 was used for GO analysis.  
```{r}
genes_up <- df$ENTREZID[df$FDR < FDR_cutoff & df$FC>FC_cutoff]

ensembl = useMart("ensembl",dataset=organism.biomart)
background_genelist = getBM(attributes=c('entrezgene_id'), mart = ensembl)[,1]
background_genelist = as.character(background_genelist)
```

#Add Enrichment Score function, Enrichment = Gene Ratio/Background Ratio
```{r}
add_enrichment <- function(go_data){
  tmp <- separate(go_data@result, GeneRatio,into= c("Gene.Ratio.x","Gene.Ratio.y"), sep="/")
  tmp <- separate(tmp, BgRatio,into= c("BgRatio.x","BgRatio.y"), sep="/")
  tmp$Enrichment <- (as.numeric(tmp$Gene.Ratio.x)/as.numeric(tmp$Gene.Ratio.y))/
    (as.numeric(tmp$BgRatio.x)/as.numeric(tmp$BgRatio.y))
  go_data@result$Enrichment <- tmp$Enrichment
  return(go_data)}
  
```

#perform GO analysis
```{r}
ego_up <- enrichGO(gene       = genes_up,
                     universe      = background_genelist,
                     OrgDb         = organism.GO,
                     ont           = ontology,
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.05,
                     qvalueCutoff  = 0.1,
                     readable      = TRUE)

```
#add enrichment score
```{r}
ego_up <- add_enrichment(ego_up)
```
#Summary for GO terms
```{r}
go_res_up <- ego_up@result[,c(2,3,6,10)]

kable(go_res_up[1:50,],caption = "GO terms for All Genes") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "300px")

write.csv(ego_up@result, file = "C:/Users/user/OneDrive/Documents/R/GO analysis/GO high confidence all.csv", sep="\t",row.names=T,col.names=T, quote=F)
```

#Summary for all ontologies
```{r}
ego_up_all <- enrichGO(gene   = genes_up,
                       universe      = background_genelist,
                       OrgDb         = organism.GO,
                       ont           = "all",
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.1,
                       qvalueCutoff  = 0.05,
                       readable      = TRUE)
tmp <- df
tmp <- tmp[,c("FC", "ENTREZID")]

#Find duplicate rows and calculate average
tmp <- aggregate(FC~ENTREZID, data = tmp, FUN = mean)

#Order by CS
tmp <- tmp[order(-tmp$FC),]

#Make Gene list
geneList <- tmp$FC
names(geneList) <- tmp$ENTREZID

```
#Plot top GO terms - different types of visualisations
```{r}
barplot(ego_up, x="Enrichment", showCategory=20,orderBy="p.adjust")+
  ggtitle("High confidence Genes")

dotplot(ego_up, x="Enrichment", showCategory=20,
        orderBy="p.adjust")+ ggtitle("High confidence Genes")

#by ontology
dotplot(ego_up_all, split="ONTOLOGY") + facet_grid(ONTOLOGY~., scale="free")+ggtitle("High confidence Genes")

```
#GO Gene-Concept Network plot
```{r}
ego_up <- simplify(ego_up)
cnetplot(ego_up, foldChange=geneList, colorEdge = TRUE)
cnetplot(ego_up, foldChange=geneList, circular = TRUE, colorEdge = TRUE)
```
#Heatmap
```{r}
#Heatmap
heatplot(ego_up, foldChange=geneList)
```

