---
title: "Visualisation and comparison of ChIP peaks between different datasets"
author: "Cosette Schupfer"
date: "04/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages

```{r, echo=FALSE, warning=FALSE}
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(clusterProfiler)
library(AnnotationDbi)
library(EnsDb.Mmusculus.v79)
library(magrittr)
library(dplyr)
library(org.Mm.eg.db)
library(UpSetR)
library(ggplot2)
library(ggimage)
```

#assgn database used for annotation (mm10 is used here specifically for the genome for mus musculus)
```{r}
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
```

#load datafiles (input: bed files from MACS output)
```{r}
#Get files
files <-getSampleFiles()
print(files)

peak <- readPeakFile(files[[5]])
```
#Convert to genomic ranges
GRanges is a package optimised for storing storing and manipulating genomic intervals along a genome. The code below also allows naming of datasets. 
```{r}
peak1 = GenomicRanges::GRangesList(GFoxg1_E14.5C1=readPeakFile(files[[3]]),
                                GFoxg1_E14.5C2=readPeakFile(files[[4]]),
                                BFoxg1_MNS=readPeakFile(files[[1]]),
                                CFoxg1_E15.5C=readPeakFile(files[[2]]),
                                SPax6_E12.5F=readPeakFile(files[[5]]))
                        
```
#Heatmap of distribution of peaks relative to TSS. 
Commonly, annotation of ChIP-peaks to likely regulatory genes is based on the assumption that transcription factor regulation occurs through binding to a cis-regulatory sequence close to the transcription start site of a gene. It is important to note that gene regulation often follows a more complex regulatory landscape. Here, we define the TSS region as 3000 pb upstream and downstream of the TSS of a gene. The heatmap gives a global evaluation of the density of peaks within the TSS region. Different colours were assigned to distinguish between datasets.  
```{r}
peakHeatmap(files[1:5], TxDb=txdb, upstream=3000, downstream=3000, color = c("magenta", "chocolate1", "seagreen2", "deeppink3", "mediumblue"))
```
#Average profile of ChIP peaks within TSS region 
Similar to the heatmap visualisation, this allows comparison of the density of peaks near the TSS between datasets. 
```{r}
plotAvgProf2(files[1:5], TxDb=txdb, upstream=3000, downstream=3000,
             xlab="Genomic Region (5'->3')", ylab = "Read Count Frequency") +  
             scale_colour_manual(values = c("magenta", "chocolate1", "seagreen2", "deeppink3", "mediumblue"))
```
#Include confidence intervals
Optionally, you may include confidence intervals calculcated through the bootstrapping method
```{r}
plotAvgProf2(files[1:5], TxDb=txdb, upstream=3000, downstream=3000, conf = 0.95, resample = 1000)+
  scale_fill_manual(values = c("magenta", "chocolate1", "seagreen2", "deeppink3", "mediumblue"))
```

